---
title: "Setting up the Stripers"
output: html_notebook
---

We have added a third lane in there, so I want to add them into the units file.
In fact, though, I don't need to add them in.  There is nothing like variant 
calling that requires doing things jointly, so, I will just make a separate units
file with them all in there, and then make a new config for them, too.  

The fastqs are at:
```
/home2/nt246_0001/eca64/striped_bass/mega-non-model-wgs-snakeflow/data
```

I got the list of them and their sizes with
```sh
(base) [eca64@cbsulogin mega-non-model-wgs-snakeflow]$ pwd
/home2/nt246_0001/eca64/striped_bass/mega-non-model-wgs-snakeflow

du   data/*/*.fastq.gz
```
And I copied that into:
```
/Users/eriq/Documents/git-repos/mega-non-model-wgs-snakeflow/example-configs/Cornell-Stripers/setup/file_listing.tsv
```

Now we can process that with R.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

f1 <- read_tsv("file_listing-Lane3.tsv") %>%
  extract(
    fq,
    into = c("junk", "sample_id", "lane", "read"),
    regex = "^data/Lane3/(B_...|B)_(.*)_10484832_237LGHLT3_L([1-9])_R([12])\\.fastq\\.gz$",
    convert = TRUE,
    remove = FALSE
  )
```

Those appear to have all parsed out.  It looks like 220 fish.

One sample had very few reads (`2015R_DCA01_04`) but we will leave
those in there and just get reports on them from fastp.  There were no samples
that looked like negative controls.


We are close to ready to go, but I want to get the barcodes and the library name
in there. I am just going to call the library the "plate", which data is available
online, but which googlesheets4 can't access for some reason.  So I just downloaded it
as a CSV.


Now we should be good to go...
```{r}

seq_meta <- read_csv("library_prep_meta.csv")

sm2 <- seq_meta %>%
  mutate(
    library = str_replace_all(`Seq plate`, "[ a-z]*", ""),
    barcode = str_c(i7, "+", i5)
  ) %>%
  select(Specimen, library, barcode)
```

Join the meta data fields on there.  When I did that the first time, I found that
there was a `KEN46` in the sequencing data that did not join to anything in the
metadata.  That is because it is called KEN46a, in the metadata, I think.  Because
there is a KEN46b.  So, this is another case of not naming the first one as it is
in the meta data I would say...

I will change its name in f1 by adding the a
```{r}
f2 <- f1 %>%
  mutate(sample_id = ifelse(sample_id == "KEN46", "KEN46a", sample_id)) %>% 
  left_join(sm2, by = join_by(sample_id == Specimen))
```

Let's see what is missing now:
```{r}
f2 %>%
  filter(is.na(barcode))
```

Nothing!  Good.  

Then we have to pivot that wider to that we have fq1 and kb1 and fq2 and kb2, and add
flowcell and sample.  We also set all of these as unit 2 because they are all
fish that were sequenced previously.  
```{r}
f3 <- f2 %>%
  pivot_wider(
    names_from = read,
    values_from = c(kb, fq),
    names_sep = ""
  ) %>%
  mutate(
    flowcell = "237LGHLT3", # got this by looking inside one of the fastqs
    sample = sample_id,
    unit = 2,
    platform = "ILLUMINA"
  )

```

Check to see how many time each sample id occurs:
```{r}
f3 %>%
  count(sample_id) %>%
  count(n)
```

They are all in there just once.  And we should check to make sure that the same sample
ids occur in the old units file from the previous run.
```{r}
old_units <- read_tsv("../config/units.tsv") %>%
  semi_join(f3, by = join_by(sample_id))

nrow(old_units)
```

That is the right number of things samples.  So, let's put columns in the right order
and then bind_rows old and new and arrange:
```{r}

final_units <- f3 %>%
  select(sample, unit, library, flowcell, platform, lane, sample_id, barcode, fq1, fq2, kb1, kb2) %>%
  bind_rows(old_units) %>%
  arrange(sample, unit)
```


Check to make sure that every sample has two units:
```{r}
final_units %>%
  count(sample) %>%
  count(n)
```
It looked good.

So, let's write this dude out.
```{r}
write_tsv(final_units, file = "../config/units-add-sequence-to-220-samples.tsv")
```

That is done.

## Making a new config

I modified config.yaml to make a new one called `config-add-seq-220.yaml`