---
title: "Setting up the Stripers"
output: html_notebook
---

The fastqs are at:
```
/home2/nt246_0001/eca64/striped_bass/mega-non-model-wgs-snakeflow/data
```

I got the list of them and their sizes with
```sh
(base) [eca64@cbsulogin mega-non-model-wgs-snakeflow]$ pwd
/home2/nt246_0001/eca64/striped_bass/mega-non-model-wgs-snakeflow

du   data/*/*.fastq.gz
```
And I copied that into:
```
/Users/eriq/Documents/git-repos/mega-non-model-wgs-snakeflow/example-configs/Cornell-Stripers/setup/file_listing.tsv
```

Now we can process that with R.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

f1 <- read_tsv("file_listing.tsv") %>%
  extract(
    fq,
    into = c("sample_id", "junk", "lane", "read"),
    regex = "^data/Lane./..._(.*)_(10483618|10483619)_2337HWLT4_L([1-9])_R([12])\\.fastq\\.gz$",
    convert = TRUE,
    remove = FALSE
  ) %>%
  select(-junk)


# some of those path names are malformed. They don't have a specifier for lane:
malformed <- f1 %>%
  filter(is.na(lane) | is.na(read))

malformed
```

That is only six files.  Four are negative controls and 2 are from a sample that
didn't get any reads.  We can just ignore those.

Now, there are also a few samples that have very few reads, but we will leave
those in there and just get reports on them from fastp.    Also, we will toss
out all the negative controls
```{r}
f2 <- f1 %>%
  filter(!(is.na(lane) | is.na(read))) %>%
  filter(!str_detect(fq, "_neg[1-9]_")) %>%
  pivot_wider(names_from = read, values_from = c(fq, kb), names_sep = "") %>%
  mutate(
    platform = "ILLUMINA",
    flowcell = "2337HWLT4",
    unit = 1, 
  )

```

We are close to ready to go, but I want to get the barcodes and the library name
in there. I am just going to call the library the "plate", which data is available
online, but which googlesheets4 can't access for some reason.  So I just downloaded it
as a CSV.

I discovered that `SJR87_L1` was named `SJR87-L1` in the meta data, so I changed that
in the CSV. Also, the letters on the possibly duplicated versions of the same
individuals were not in the meta data, so I added those in the order in which the
samples appeared (multiple times) in the CSV.


Namely, I had to do this with:
```
"KEN46a"     "ROA_SB128b" "ROA_SB160b" "ROA_SB138b" "HUD14_E36b" "KEN46b"     "KEN46c"
```
Harmony gave me a heads up about these samples...

But, it looks like there are also two version of `RAP14_03`, so I renamed the second
occurrence of it with a b suffix in the meta-data CSV, and I also will change
it in R here for the Snakemake units file:
```{r}
f3 <- f2 %>%
  mutate(
    idx = 1:n(),
    sample_id = ifelse(idx == 310 & sample_id == "RAP14_03", "RAP14_03b", sample_id),
    sample = sample_id
  ) %>%
  select(-idx)
```

Now we should be good to go...
```{r}

seq_meta <- read_csv("library_prep_meta.csv")

sm2 <- seq_meta %>%
  mutate(
    library = str_replace_all(`Seq plate`, "[ a-z]*", ""),
    barcode = str_c(i7, "+", i5)
  ) %>%
  select(Specimen, library, barcode)

# now, join that
f4 <- f3 %>%
  left_join(sm2, by = join_by(sample == Specimen))
```

And finally, we just need to get the columns in the correct order:
```{r}
final_units <- f4 %>%
  select(sample, unit, library, flowcell, platform, lane, sample_id, barcode, fq1, fq2, kb1, kb2)
```

That gives us 546 samples.   So, who is missing from the seq_meta?
```{r}
seq_meta %>% anti_join(final_units, by = join_by(Specimen == sample))

setdiff(seq_meta$Specimen, final_units$sample)
```

That is cool, we are just missing:
```
"CB208" "neg1"  "neg2"  "neg3"  "neg4"  "neg5"  "neg6"  "neg7"  "neg8"  "neg9"
```
And Harmony told me that CB208 had failed.

So, let's write this dude out.
```{r}
write_tsv(final_units, file = "../config/units.tsv")
```

That is done.

## Getting the chromosomes and scaffold groups...

This we actually don't need to do for the parts of the workflow we are doing...

## Testing

I am going to run through one of them all the way to the overlap-clipped ones.
By the way.  That is where we will toss the improper pairs.

Here is the one we will do:
```
results/bqsr-round-0/overlap_clipped/HUD12_26.bam
```
That one is about 187 Mb per read, so it is smaller than most others.  

I worked all the kinks out on that and then got a jdb slurm profile up and working.

I did a dry-run with this:
```sh
snakemake -np  results/bqsr-round-0/qc/multiqc.html  dest_overlap_clipped_bams  --configfile example-configs/Cornell-Stripers/config/config.yaml --profile smk-9-slurm/jdb/biohpc
```
And the dry-run report looks like:
```
Shell command: None
Job stats:
job                          count
-------------------------  -------
clip_overlaps                  543
dest_overlap_clipped_bams        1
map_reads                      543
mark_duplicates                543
multiqc_dir                      1
samtools_stats                 546
trim_reads_pe                  543
total                         2720
```

So, I am going to give it a big ol' whirl now.

That ran pretty well, but many of the MarkDuplicates jobs failed.  I am guessing
the ran out of memory or time.

I did a dry-run, and there are about 213 remaining for mark duplicates.  I gave that
rule 8192 Mb per cpu and will let it rip again...
```

```