---
title: 'Set Up the "To Be Assigned" Striped Bass'
output: html_notebook
---



There are a lot of these fish.  Fun.  It will be a big run. The data are in:
```
/home2/nt246_0001/nt246/striped_bass/samples_for_assignment/
```
and I make a symlink to that in
```
/home2/nt246_0001/eca64/striped_bass/mega-non-model-wgs-snakeflow/data
```


We get a file listing of the fastqs with:
```sh
ls -l data/samples_for_assignment/*.fastq.gz | awk 'BEGIN {printf("kb\tfq\n");} {printf("%.0f\t%s\n", $5/1000, $NF);}'
```
and I copied that into:
```
example-configs/Cornell-Stripers/setup/to-be-assigned-fastq-listing.txt
```

Then, I also got the library prep info from:
```
https://docs.google.com/spreadsheets/d/1FbrdFyWuhZhJZvzbwceu3c0Z9p5RotsS8JKIwP-rrOM/edit?gid=182325693#gid=182325693
```
and downloaded that to:
```
example-configs/Cornell-Stripers/setup/to-be-assigned-hackflex-libraries.csv
```

Note that I am going to add the negative controls and the topunkowns in the
analysis.



## Making the units file

```{r}
library(tidyverse)
fastqs <- read_tsv("to-be-assigned-fastq-listing.txt") %>%
  extract(
    fq,
    into = c("sample_id", "lane", "read"),
    regex = "^data/samples_for_assignment/A_..._(.*)_10484831_237LGHLT3_L([1-9])_R([12])\\.fastq\\.gz$",
    convert = TRUE,
    remove = FALSE
  )
```

Check on any that did not get parsed
```{r}
fastqs %>%
  filter(is.na(sample_id)) %>% View()
```

It is just the top Unknown ones.  So, we can fix that
```{r}
fastqs[is.na(fastqs$sample_id), "sample_id"] <- "TopUnknown"
fastqs[is.na(fastqs$lane), "lane"] <- 5L
fastqs[is.na(fastqs$read), "read"] <- c(1L, 2L)
```


We can pivot this wider and then attach meta data:
```{r}
fwide <- fastqs %>%
  pivot_wider(
    names_from = read,
    values_from = c(kb, fq),
    names_sep = ""
  )
```

Then join the library metadata.  Note that I edited the library file by removing a
space in `neg 6` and also removing some barcode-check columns that were not attached
to any samples.

```{r}
libmeta <- read_csv("to-be-assigned-hackflex-libraries.csv") %>%
  mutate(
    library = str_replace_all(plate, " ", "-"),
    barcode = paste(i7, i5, sep = "+")
  ) %>%
  rename(sample_id = `Specimen ID`) %>%
  select(sample_id, library, barcode)

joined <- fwide %>%
  full_join(libmeta, by = join_by(sample_id))
```

Now, we need to find any joining failures:
```{r}
joined %>%
  filter(is.na(lane) | is.na(barcode)) %>% View()
```

So, the only one that doesn't have something going on is the TopUnknown.  We can
add the barcode and the Library to that.  It didn't formally come from a particular
library, so we just say it came from the TopUnknown library.
```{r}
joined[is.na(joined$library), "library"] <- "TopUnkLib"
joined[is.na(joined$barcode), "barcode"] <- "GCTCATCA+TGCGGAAC"
```

Now we just need to get a few more columns and put them in the right order.
```{r}
units <- joined %>%
  mutate(
    flowcell = "237LGHLT3",
    platform = "ILLUMINA",
    sample = sample_id,
    unit = 1
  ) %>%
  select(sample, unit, library, flowcell, platform, lane, sample_id, barcode, fq1, fq2, kb1, kb2)

View(units)

write_tsv(units, file = "../config/units-Canadian-fish-to-be-assigned.tsv")
```


## Launching the runs:

Dry run with this:

```sh
snakemake -np  results/bqsr-round-0/qc/multiqc.html  dest_overlap_clipped_bams  --configfile example-configs/Cornell-Stripers/config/config-Canadian-fish-to-be-assigned.yaml --profile smk-9-slurm/jdb/biohpc
```
Gave me this:
```sh
Shell command: None
Job stats:
job                          count
-------------------------  -------
clip_overlaps                  857
dest_overlap_clipped_bams        1
map_reads                      857
mark_duplicates                857
multiqc_dir                      1
samtools_stats                 857
trim_reads_pe                  857
total                         4287
```
That looks spot on! Launching it now.

```sh
snakemake -p  results/bqsr-round-0/qc/multiqc.html  dest_overlap_clipped_bams  --configfile example-configs/Cornell-Stripers/config/config-Canadian-fish-to-be-assigned.yaml --profile smk-9-slurm/jdb/biohpc
```